---
title: "Final Project"
author: "Bailey Vande Griend"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = TRUE)
```
<font size="6"> How has Baseball Changed Recently? </font>

It's no secret that baseball's popularity has been in question recently. Some say its a dying sport, while others insist it is only stagnant and still has room to grow. Regardless, many efforts have been made to make baseball more exciting to watch. One such change involved changing the league standard for the baseball. This was done in 2017, and more recently in 2019. The question is, has this actually affected the sport? To answer this question, we will be looking at hitting stats from the last 4 years to see if there are any differences.

```{r}
setwd("/Users/baile/OneDrive/Git/Stat-344")
library(readxl)
Bball2016 <- read_excel("2016Stats.xlsx")
Bball2017 <- read_excel("2017Stats.xlsx")
Bball2018 <- read_excel("2018Stats.xlsx")
Bball2019 <- read_excel("2019Stats.xlsx")
```

