---
title: "HW2"
author: "Bailey Vande Griend"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```
**Question 1**

```{r}
setwd("/Users/baile/OneDrive/Stat")
messyData <- read.csv("MessyData", header = TRUE, stringsAsFactors = FALSE)
messyDataOriginal <- data.frame(read.csv("MessyData"))
```
Some of the columns are hard to understand. It is hard to tell what Q1, Q2, Q3
are describing. Also, weight does not have a unit of measurement. There are plenty 
of missing data points as well. There also appear to be some mistakes in the names,
as one of them seems to be entered twice, just incorrectly one of the times and
another is in all caps. The first half also seems to be in alphabetical order, but 
then afterwards is not. In Q3, some data points are simplified to "y" and "n". while
others are written out to be "yes" and "no."

**Question 2**

"Wedel" appears in both the First and Last category, and the last name Brown appears
in Last twice

**Question 3**

```{r}
extraComments <- read.csv("Extra.csv", header = TRUE, stringsAsFactors = FALSE)
sampVect <- match(extraComments$Last, messyData$Last)
```
These last names do not exist in the original data so we delete them
```{r}
sampVect <- sampVect[-3]
sampVect <- sampVect[-3]
sampVect <- sampVect[-5]
extraComments <- extraComments[-3,]
extraComments <- extraComments[-3,]
extraComments <- extraComments[-5,]
for (i in 1:length(extraComments$comments))
messyData[sampVect[i],8] <- extraComments$comments[i]
```

**Question 4**

```{r}
messyData$Q3 <- sub("y", "Y", messyData$Q3)
messyData$Q3 <- sub("Yes", "Y", messyData$Q3)
messyData$Q3 <- sub("n", "N", messyData$Q3)
```

**Question 5**

```{r}
messyData$Last <- gsub("-.*","",messyData$Last)
```

**Question 6**

```{r}
afterDecimal <- gsub(".*\\.","", as.character.numeric_version(messyData$weight))
messyData$weight <- gsub("\\..*", "", messyData$weight)
messyData["WeightDec"] <- afterDecimal
```

**Question 7**

Delete pointless column and the name that was entered incorrectly
```{r}
messyData <- messyData[-4,]
messyData <- messyData[-6]
```
Fix capitalization of one name
```{r}
messyData$First <-  sub("CLARA", "Clara", messyData$First)
messyData$Last <-  sub("POMPER", "Pomper", messyData$Last)
```
Change the titles of the columns so they are all capitalized
```{r}
colnames(messyData)[3] <- "Age"
colnames(messyData)[4] <- "Weight(kg)"
colnames(messyData)[7] <- "Comments"
```

**Question 8**

```{r}
setwd("/Users/baile/OneDrive/Stat/IowaPovLayouts")
layouts <- list()
layoutDirectory <- dir(pattern = "\\d+-estimate-layout.txt")
for (i in 1:length(layoutDirectory)){
  layouts[[i]] <- readLines(layoutDirectory[i])
}

setwd("/Users/baile/OneDrive/Stat/IowaPov")
datasets <- list()
dataDirectory <- dir(pattern = "Iowa\\d{4}.txt")
for (i in 1:length(dataDirectory)){
  datasets[[i]] <- readLines(dataDirectory[i])
}
countyStartIndex <- list()
countyEndIndex <- list()
county <- list()
countyWidths <- list()

for(i in 1:length(layouts)){
  county[i] <- grep("[0-9]", layouts[[i]][grep("State or county name", layouts[[i]])], value = TRUE)
  county[i] <- gsub("[^0-9-]","",county[i])
  
  countyStartIndex[i] <- as.numeric(gsub("-\\d+","", county[i]))
  countyEndIndex[i] <- as.numeric(gsub("\\d+-","", county[i]))
}
countyStartIndex <- unlist(countyStartIndex)
countyEndIndex <- unlist(countyEndIndex)
for (i in 1:length(countyStartIndex)){
  countyWidths[i] <- countyEndIndex[i]-countyStartIndex[i]
}

povertyStartIndex <- list()
povertyEndIndex <- list()
poverty <- list()
povertyWidths <- list()

for (i in 1:length(layouts)){
  poverty[i] <- grep("[0-9]", layouts[[i]][grep("Estimated percent of people of all ages in poverty", layouts[[i]])], value = TRUE)
  poverty[i] <- gsub("[^0-9-]", "", poverty[i])
  
  povertyStartIndex[i] <- as.numeric(gsub("-\\d+","",poverty[i]))
  povertyEndIndex[i] <- as.numeric(gsub("\\d+-", "", poverty[i]))
}

povertyStartIndex <- unlist(povertyStartIndex)
povertyEndIndex <- unlist(povertyEndIndex)
for (i in 1:length(povertyStartIndex)){
  povertyWidths[i] <- povertyEndIndex[i]-povertyStartIndex[i]
}

LowerStartIndex <- list()
LowerEndIndex <- list()
Lower <- list()
LowerWidths <- list()

for (i in 1:length(layouts)){
  LowerMatches <- grep("90% confidence interval lower bound of", layouts[[i]])
  Lower[i] <- grep("[0-9]", layouts[[i]][LowerMatches[2]], value = TRUE)
  Lower[i] <- gsub("[^0-9-]", "", Lower[i])
  Lower[i] <- gsub("90", "", Lower[i])
  
  LowerStartIndex[i] <- as.numeric(gsub("-\\d+", "", Lower[i]))
  LowerEndIndex[i] <- as.numeric(gsub("\\d+-", "", Lower[i]))
}

LowerStartIndex <- unlist(LowerStartIndex)
LowerEndIndex <- unlist(LowerEndIndex)
for (i in 1:length(LowerStartIndex)){
  LowerWidths[i] <- LowerEndIndex[i]-LowerStartIndex[i]
}

widths <- list()
for (i in 1:length(layouts)){
  widths[[i]] <- c(as.numeric(povertyStartIndex[i])-1, as.numeric(povertyWidths[i])+2,
    as.numeric(LowerWidths[i])+2, as.numeric(countyStartIndex[i])-(as.numeric(LowerEndIndex[i])+2), 
    as.numeric(countyWidths[i])+2,3)
}
for (i in 1:length(dataDirectory)){
  datasets[[i]] <- read.fwf(dataDirectory[i], widths[[i]])
  datasets[[i]] <- datasets[[i]][c(2,3,5)]
  colnames(datasets[[i]])[1] <- "PovertyPercent"
  colnames(datasets[[i]])[3] <- "CountyOrStateName"
  for (j in 1:length(datasets[[i]]$PovertyPercent)){
  datasets[[i]]["PercentMOE"] <- datasets[[i]]$PovertyPercent-datasets[[i]]$V3
  }
  datasets[[i]]["Year"] <- i+1996
  datasets[[i]] <- datasets[[i]][c(1,3,4,5)]
}
```

This is the combined data frames, its meant to be read as a collection of 
separate data frames rather than one large data frame
```{r}
DFdatasets <- data.frame(datasets)
```